### 2023-12-20 manuscript.tex
. [ci skip]
This build is based on
https://github.com/MMV-Lab/im2im-paper/commit/9e8e8c612f56fb5296c1a4e6be2cb6cd579e4008.

This commit was created by the following CI build and job:
https://github.com/MMV-Lab/im2im-paper/commit/9e8e8c612f56fb5296c1a4e6be2cb6cd579e4008/checks
https://github.com/MMV-Lab/im2im-paper/actions/runs/7279129324

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/3587e45ed5b57a2b3dda176e1e9881404c079911/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/3587e45ed5b57a2b3dda176e1e9881404c079911}{MMV-Lab/im2im-paper@3587e45}
on December 20, 2023.

@@

For computational biomedical researchers, \emph{MMV\_Im2Im} provides a starting point for developing new biomedical image analysis or machine learning algorithms, where they can either reuse the code in this package or fork and extend this package to facilitate the development of new methods. Experimental biomedical researchers can benefit from this work by gaining a comprehensive view of the image-to-image transformation concept through diversified examples and use cases. We hope this work can give the community inspirations on how DL-based image-to-image transformation can be integrated into the assay development process, enabling new biomedical studies that cannot be done only with traditional experimental assays. To help researchers get started, we have provided source code, documentation, and tutorials for \emph{MMV\_Im2Im} at https://github.com/MMV-Lab/mmv\_im2im under MIT license.

@@

With the rapid advancements in the fields of machine learning (ML) and computer vision, computers can now transform images into new forms, enabling better visualization {[}\protect\hyperlink{ref-1O0bopKD}{1}{]}, better animation {[}\protect\hyperlink{ref-LxlUp436}{2}{]} and better information extraction {[}\protect\hyperlink{ref-11chATuF4}{3}{]} with unprecedented and continuously growing accuracy and efficiency compared to conventional digital image processing. These techniques have recently been adapted for bioimaging applications and have revolutionized image-based biomedical research {[}\protect\hyperlink{ref-Yq8wZ6hc}{4},\protect\hyperlink{ref-WwenuBHa}{5},\protect\hyperlink{ref-wcCVn8av}{6},\protect\hyperlink{ref-xPgDok51}{7}{]}. In principle, these techniques and applications can be formulated as a general image-to-image transformation problem, as depicted in the central panel in Figure \ref{fig:overview}. Deep neural networks are trained to perceive the information from the source image(s) and reconstruct the learned knowledge from source images(s) in the form of a new image(s) of the target type. The source and target images can be real or simulated microscopy images, segmentation masks, or their combinations, as exemplified in Figure \ref{fig:overview}. Since these underlying methods share the same essential spirit, a natural question arises: is it possible to develop a single generic codebase for deep learning (DL) based image-to-image transformation applicable to various biomedical studies?

In this paper, we introduce \emph{MMV\_Im2Im} an open-source microscopy machine vision (MMV) toolbox for image-to-image transformation and demonstrate its applications in over 10 biomedical tasks of various types by performing more than 30 experiments. Currently, \emph{MMV\_Im2Im} supports handling 2D\textasciitilde5D microscopy images for supervised image-to-image translation (e.g., labelfree determination {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}, imaging modality transformation {[}\protect\hyperlink{ref-WwenuBHa}{5},\protect\hyperlink{ref-UEBDZ3tI}{8}{]}), supervised image restoration {[}\protect\hyperlink{ref-wcCVn8av}{6}{]}, supervised semantic segmentation {[}\protect\hyperlink{ref-TutLhFSz}{9}{]}, supervised instance segmentation {[}\protect\hyperlink{ref-K2ugNcVa}{10},\protect\hyperlink{ref-QmYuUQ5K}{11}{]}, unsupervised semantic segmentation {[}\protect\hyperlink{ref-RuFP3CS3}{12}{]}, unsupervised image to image translation and synthetization {[}\protect\hyperlink{ref-6wtIu4QY}{13}{]}. The toolbox will continuously grow with more and more methods, such as self-supervised learning based methods, ideally also with contributions from the open-source community.

@@

Our package \emph{MMV\_Im2Im} employs pytorch-lightning {[}\protect\hyperlink{ref-YbvSvdyB}{14}{]} as the core in the backend, which offers numerous benefits, such as readability, flexibility, simplicity and reusability. First of all, have you ever had the moment when you wanted to extend someone's open-source code to suit your special ML needs, but found it so difficult to figure out where and how to extend, especially for complex methods? Or, have you ever encountered the situation where you want to compare the methods and code from two different papers, even solving the same problem, e.g.~semantic segmentation, but not quite easy to grasp quickly since the two repositories are implemented in very different ways? It is not rare that even different researchers from the same group may implement similar methods in very different manners. This is not only a barrier for other people to learn and re-use the open-source code, but also poses challenges for developers in maintenance, further development, and interoperability among different packages. We follow the pytorch-lightning framework and carefully design a universal boilerplate for image-to-image transformation for biomedical applications, where the implementation of all the methods share the same modularized code structure. For all PyTorch users, this greatly lowers the learning curve for people to read and understand the code, and makes implementing new methods or extending existing methods simple and fast, at least from an engineering perspective.

Moreover, as ML scientists, have you ever been overwhelmed by different training tricks for different methods or been curious about if certain state-of-the-art training methods can boost the performance of existing models? With the pytorch-lightning backend, \emph{MMV\_Im2Im} allows you to enjoy different state-of-the-art ML training engineering techniques without changing the code, e.g., stochastic weight averaging {[}\protect\hyperlink{ref-qzeQFRn9}{15}{]}, single precision training, automatic batch size determination, different optimizers, different learning rate schedulers, easy deployment on different devices, distributed training on multi-GPU (even multi-node), logging with common loggers such as Tensorboard, etc.. In short, with the pytorch-lightning based universal boilerplate, bioimaging researchers can really focus on research and develop novel methods for their biomedical applications, without worrying about the ML engineering works (which are usually lacking in non-computer-science labs).

@@

The original idea of a general toolbox actually stemmed from the OpenMMLab project (\url{https://openmmlab.com/}), which provides generic codebases for a wide range of computer vision research topics. For instance, \emph{MMSegmentation} {[}\protect\hyperlink{ref-EcizztLg}{16}{]} is an open source toolbox for semantic segmentation, supporting unified benchmarking and state-of-the-art models ready to use out-of-box. It has become one of most widely used codebase for research in semantic segmentation (2.3K forks and 6.5K stars on Github as of September 29, 2023). This inspires us to develop \emph{MMV\_Im2Im} to facilitate research in image-to-image transformation with a special focus on biomedical applications.

First of all, different from general computer vision datasets, such as ImageNet {[}\protect\hyperlink{ref-lt4BNUoG}{17}{]}, where the images are usually small 2D RGB images (e.g., 3 x 256 x 256 pixels), biomedical applications usually involves large-scale high dimensional data (e.g., 500 images of 4 x 128 x 2048 x 2048 voxels). To deal with this issue, we employ the PersistentDataset in MONAI {[}\protect\hyperlink{ref-UU62HYC6}{18}{]} with partial loading and sampling support, as well as delayed image reading powered by aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{19}{]} as default (configurable if another dataloader is preferred). As a result, in our stress test, training a 3D nuclei instance segmentation model with more than 125,000 3D images can be conducted efficiently in a day, even with limited resources.

@@

Third, the toolbox pre-packs common functionalities specific to microscopy images. For example, we incorporate the special image normalization method introduced in {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}, where only the middle chunk along the Z dimension of 3D microscopy images will be used for calculating the mean and standard deviation of image intensity for standard normalization. Also, 3D light microscopy images are usually anisotropic, i.e., much lower resolution along Z than XY dimension. So, we adopt the anisotropic variation of UNet as proposed in {[}\protect\hyperlink{ref-jM3v1UjQ}{20}{]}.

Finally, to deploy the model in production, a model trained on small 3D patches sometimes needs to be applied not only on much larger images. Combining the efficient data handling of aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{19}{]} and the sliding window inference with gaussian weighted blending, the toolbox can yield efficient inference without visible stitching artifacts in production.

All in all, the \emph{MMV\_Im2Im} toolbox stands on the shoulders of many giants in the open-source software and ML engineering communities (pytorch-lightning, MONAI, aicsimageio, etc.) and is systematically designed for image-to-image transformation R\&D for biomedical applications. The source code of \emph{MMV\_Im2Im} is available at \url{https://github.com/MMV-Lab/mmv_im2im}. This manuscript is generated with the open-source package Manubot {[}\protect\hyperlink{ref-YuJbg3zO}{21}{]}. The manuscript source code is available at \url{https://github.com/MMV-Lab/im2im-paper}.

@@

The labelfree method refers to a DL method that can predict fluorescent images directly from transmitted light brightfield images {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. Compared to brightfield images, fluorescent images can resolve subcellular structures in living cells at high resolution but with the cost of expensive and slow procedures and high phototoxicity. The labelfree method provides a new perspective in assay development to conduct integrated computational analysis of multiple organelles only with a single brightfield image acquisition. In our first demonstration, we applied \emph{MMV\_Im2Im} to build 2D/3D models that can predict fluorescent images of nuclear structures from brightfield images. For 3D models, we also compared (1) different image normalization methods, (2) different network backbones, and (3) different types of models.

It should be noted that while we recognize the importance of systematically evaluating the predictions, such an analysis falls outside the scope of this paper. We argue that an appropriate evaluation methodology should depend on specific downstream quantitative analysis goals (e.g., {[}\protect\hyperlink{ref-C2iqR6xE}{24}{]}). For example, if our aim is to quantify the size of nucleoli, we must compare the segmentation derived from real nucleoli signals to that of the predicted nucleoli segmentation, ensuring that measurements from both are consistent. Alternatively, if the goal is to localize the nucleoli roughly within the cell, Pearson correlation may be a more appropriate metric. In this work, we concentrate on visual inspection, using Pearson correlation and structural similarity as a rough quantitative reference. Our intent is to demonstrate the utility of our \emph{MMV\_Im2Im} package, and leave appropriate evaluations to users in their specific problems in real studies.

\emph{2D Labelfree:} We started with a simple problem using 2D images from the HeLa ``Kyoto'' cells dataset {[}\protect\hyperlink{ref-xv2VIyRP}{25}{]}. For all images, we took the brightfield channel and the mCherry-H2B channel out of the multi-channel timelapse movies. 2D images were acquired at 20x with 0.8 N.A. and then downscaled by 4 (pixel size: 0.299 nm x 0.299 nm). Example predictions can be found in Figure \ref{fig:labelfree_comparison}-A. We compared a basic UNet model {[}\protect\hyperlink{ref-TutLhFSz}{9}{]} and a 2D version of the fnet model in {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. The fnet model achieved slightly more accurate predictions than the basic UNet, as seen in Figure \ref{fig:labelfree_comparison}-A.

\emph{3D Labelfree:} We tested with 3D images from the hiPSC single cell image dataset {[}\protect\hyperlink{ref-5sGcmDuy}{26}{]}. Specifically, we extracted the brightfield channel and the structure channel from the full field-of-view (FOV) multi-channel images, from the HIST1H2BJ, FBL, NPM1, LMNB1 cell lines, so as to predict from one brightfield image various nuclear structures, histones, nucleoli (dense fibrillar component via fibrillarin), nucleoli (granular component via nucleophosmin), and nuclear envelope, respectively. Images were acquired at 100x with 1.25 NA (voxel size: 0.108 micron x 0.108 micron x 0.29 micron).

@@

We conducted three groups of comparisons (see results in Figure \ref{fig:labelfree_comparison}-B). First, we compared three different image normalization methods for 3D images: percentile normalization, standard normalization, center normalization {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. Percentile normalization refers to cutting the intensity out of the range of {[}0.5, 99.5{]} percentile of the image intensity and then rescale the values to the range of {[}-1, 1{]}, while the standard normalization is simply subtracting mean intensity and then divided by the standard deviation of all pixel intensities. Center normalization is similar to standard normalization, but the statistics are calculated only around the center along the Z-axis {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. One could easily test different percentiles or rescaling to {[}0, 1{]} instead of {[}-1, 1{]}. Qualitatively, we found center normalization slightly more accurate and more robust than the other two (see first row of Figure \ref{fig:labelfree_all_structures}-B).

Second, we compared different network backbone architectures, including the original fnet model {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}, an enhanced UNet {[}\protect\hyperlink{ref-M7480NLD}{27}{]}, the attention UNet {[}\protect\hyperlink{ref-OCow1hly}{28}{]}, two transformer-based models, SwinUNETR {[}\protect\hyperlink{ref-ZWL3IrVc}{29}{]} and UNETR{[}\protect\hyperlink{ref-XCKUntOB}{30}{]} (all with center normalization). Inspecting the predictions on a holdout validation set suggested that fnet achieved the best performance.

@@

Segmentation is a common image processing task, and can be considered as a special type of image-to-image transformation, where the generated images are segmentation masks. DL-based methods have achieved huge success in semantic segmentation in biomedical images. In this example, we demonstrated \emph{MMV\_Im2Im} on a pathology application to segment glands from hematoxylin and eosin (H\&E) stained tissue images from the 2015 Gland Segmentation challenge {[}\protect\hyperlink{ref-45Sirz1X}{31},\protect\hyperlink{ref-XAffSYIR}{32}{]}. Stain normalization is an important pre-processing step in order to develop models robust to stain variation and tissue variations. \emph{MMV\_Im2Im} included a classic stain normalization method {[}\protect\hyperlink{ref-tQhnZyjK}{33}{]} as a pre-processing step. The effect of stain normalization can be observed in Figure \ref{fig:2d_gland}-A and B. We trained a simple attention UNet model {[}\protect\hyperlink{ref-OCow1hly}{28}{]}. Evaluated on the two different holdout test sets, the model achieved F1-score, 0.904 ± 0.060 and 0.861 ± 0.117 on test set A and test set B, respectively. The performance was competitive compared to the methods reported in the challenge report {[}\protect\hyperlink{ref-XAffSYIR}{32}{]}, especially with much more consistent performance across the two different test sets. Example results can be found in Figure \ref{fig:2d_gland}-C.

@@

Instance segmentation is a type of segmentation problem that goes beyond semantic segmentation. The goal is to differentiate not only between different types of objects, but also different instances of the same type of objects. Currently, the \emph{MMV\_Im2Im} package supports EmbedSeg-type models. The major benefit of EmbedSeg-type models is their agnosticism to the morphology and dimensionality of the object instances, compared to other models such as StarDist {[}\protect\hyperlink{ref-tIIG2f8K}{34},\protect\hyperlink{ref-14h90Vfg0}{35}{]}, SplineDist {[}\protect\hyperlink{ref-17Yrl6WGQ}{36}{]} and Cellpose {[}\protect\hyperlink{ref-TugPkOLy}{37}{]}. For example, different from the others, EmbedSeg-type models are even able to generate instance segmentation where each instance contains multiple connected components. Additional frameworks such as Omnipose {[}\protect\hyperlink{ref-lXzmjM5n}{38}{]} will be supported in future versions. Another mainstream category of instance segmentation methods are detection-based models, such as Mask-RCNN {[}\protect\hyperlink{ref-xi8wnibR}{39}{]}. However, these models are better suited to the detection framework rather than image-to-image transformation (see Discussion section for details).

The \emph{EmbedSeg}-type models were re-implemented according to the original paper {[}\protect\hyperlink{ref-K2ugNcVa}{10},\protect\hyperlink{ref-QmYuUQ5K}{11}{]} following the generic boilerplate in \emph{MMV\_Im2Im}, with significant improvement. First of all, following the modular design of \emph{MMV\_Im2Im}, it is flexible to use different neural network models as the backbone. For 3D anisotropic microscopy images, the original backbone ERFNet {[}\protect\hyperlink{ref-XAkgs3Nh}{40}{]} doesn't take the anisotropic dimensions into account and therefore may not perform well or even be applicable. In this scenario, it is straightforward to employ another anisotropic neural network bone, such as the anisotropic U-Net in {[}\protect\hyperlink{ref-jM3v1UjQ}{20}{]} or the anisotropic version of Dynamic U-Net in MONAI. Second, we significantly improve training strategy. The original version requires pre-cropping patches centered around each instance and pre-calculated the center images and class images. This may generate a massive amount of additional data on the disk. More importantly, such pre-cropping makes data augmentation nearly impossible, except the simple ones like flipping (otherwise, the pre-calculated centers might be wrong), and also greatly undersamples around negative cases (e.g., background). For example, we have observed that for an EmbedSeg model training only with patches centered around instances, the model may suffer from degraded performance during inference when there are a large amount of background areas without any instances. Again, following the modular design of \emph{MMV\_Im2Im}, it is now possible to do on-the-fly data augmentation and patch sampling, even weighted patch sampling. Third, our improved \emph{EmbedSeg}-type models can accept an exclusion mask so that certain parts of the images can be ignored during training. This is especially useful for partially annotated ground truth. For large images, it could be extremely time-consuming to require every single instance to be annotated. The exclusion masks can address this bottleneck. Another extension compared to the original implementation was that the \emph{MMV\_Im2Im} package made sliding windowing inference straightforward, and therefore permitted easy handling of images of any size during inference.

>>>>

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/9e8e8c612f56fb5296c1a4e6be2cb6cd579e4008/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/9e8e8c612f56fb5296c1a4e6be2cb6cd579e4008}{MMV-Lab/im2im-paper@9e8e8c6}
on December 20, 2023.

@@

For computational biomedical researchers, \emph{MMV\_Im2Im} provides a starting point for developing new biomedical image analysis or machine learning algorithms, where they can either reuse the code in this package or fork and extend this package to facilitate the development of new methods. Experimental biomedical researchers can benefit from this work by gaining a comprehensive view of the image-to-image transformation concept through diversified examples and use cases. We hope this work can give the community inspirations on how DL-based image-to-image transformation can be integrated into the assay development process, enabling new biomedical studies that cannot be done only with traditional experimental assays. To help researchers get started, we have provided source code, documentation, and tutorials for \emph{MMV\_Im2Im} at {[}\protect\hyperlink{ref-10dtMviwb}{1}{]} under MIT license.

@@

With the rapid advancements in the fields of machine learning (ML) and computer vision, computers can now transform images into new forms, enabling better visualization {[}\protect\hyperlink{ref-1O0bopKD}{2}{]}, better animation {[}\protect\hyperlink{ref-LxlUp436}{3}{]} and better information extraction {[}\protect\hyperlink{ref-11chATuF4}{4}{]} with unprecedented and continuously growing accuracy and efficiency compared to conventional digital image processing. These techniques have recently been adapted for bioimaging applications and have revolutionized image-based biomedical research {[}\protect\hyperlink{ref-Yq8wZ6hc}{5},\protect\hyperlink{ref-WwenuBHa}{6},\protect\hyperlink{ref-wcCVn8av}{7},\protect\hyperlink{ref-xPgDok51}{8}{]}. In principle, these techniques and applications can be formulated as a general image-to-image transformation problem, as depicted in the central panel in Figure \ref{fig:overview}. Deep neural networks are trained to perceive the information from the source image(s) and reconstruct the learned knowledge from source images(s) in the form of a new image(s) of the target type. The source and target images can be real or simulated microscopy images, segmentation masks, or their combinations, as exemplified in Figure \ref{fig:overview}. Since these underlying methods share the same essential spirit, a natural question arises: is it possible to develop a single generic codebase for deep learning (DL) based image-to-image transformation applicable to various biomedical studies?

In this paper, we introduce \emph{MMV\_Im2Im} an open-source microscopy machine vision (MMV) toolbox for image-to-image transformation and demonstrate its applications in over 10 biomedical tasks of various types by performing more than 30 experiments. Currently, \emph{MMV\_Im2Im} supports handling 2D\textasciitilde5D microscopy images for supervised image-to-image translation (e.g., labelfree determination {[}\protect\hyperlink{ref-Yq8wZ6hc}{5}{]}, imaging modality transformation {[}\protect\hyperlink{ref-WwenuBHa}{6},\protect\hyperlink{ref-UEBDZ3tI}{9}{]}), supervised image restoration {[}\protect\hyperlink{ref-wcCVn8av}{7}{]}, supervised semantic segmentation {[}\protect\hyperlink{ref-TutLhFSz}{10}{]}, supervised instance segmentation {[}\protect\hyperlink{ref-K2ugNcVa}{11},\protect\hyperlink{ref-QmYuUQ5K}{12}{]}, unsupervised semantic segmentation {[}\protect\hyperlink{ref-RuFP3CS3}{13}{]}, unsupervised image to image translation and synthetization {[}\protect\hyperlink{ref-6wtIu4QY}{14}{]}. The toolbox will continuously grow with more and more methods, such as self-supervised learning based methods, ideally also with contributions from the open-source community.

@@

Our package \emph{MMV\_Im2Im} employs pytorch-lightning {[}\protect\hyperlink{ref-YbvSvdyB}{15}{]} as the core in the backend, which offers numerous benefits, such as readability, flexibility, simplicity and reusability. First of all, have you ever had the moment when you wanted to extend someone's open-source code to suit your special ML needs, but found it so difficult to figure out where and how to extend, especially for complex methods? Or, have you ever encountered the situation where you want to compare the methods and code from two different papers, even solving the same problem, e.g.~semantic segmentation, but not quite easy to grasp quickly since the two repositories are implemented in very different ways? It is not rare that even different researchers from the same group may implement similar methods in very different manners. This is not only a barrier for other people to learn and re-use the open-source code, but also poses challenges for developers in maintenance, further development, and interoperability among different packages. We follow the pytorch-lightning framework and carefully design a universal boilerplate for image-to-image transformation for biomedical applications, where the implementation of all the methods share the same modularized code structure. For all PyTorch users, this greatly lowers the learning curve for people to read and understand the code, and makes implementing new methods or extending existing methods simple and fast, at least from an engineering perspective.

Moreover, as ML scientists, have you ever been overwhelmed by different training tricks for different methods or been curious about if certain state-of-the-art training methods can boost the performance of existing models? With the pytorch-lightning backend, \emph{MMV\_Im2Im} allows you to enjoy different state-of-the-art ML training engineering techniques without changing the code, e.g., stochastic weight averaging {[}\protect\hyperlink{ref-qzeQFRn9}{16}{]}, single precision training, automatic batch size determination, different optimizers, different learning rate schedulers, easy deployment on different devices, distributed training on multi-GPU (even multi-node), logging with common loggers such as Tensorboard, etc.. In short, with the pytorch-lightning based universal boilerplate, bioimaging researchers can really focus on research and develop novel methods for their biomedical applications, without worrying about the ML engineering works (which are usually lacking in non-computer-science labs).

@@

The original idea of a general toolbox actually stemmed from the OpenMMLab project (\url{https://openmmlab.com/}), which provides generic codebases for a wide range of computer vision research topics. For instance, \emph{MMSegmentation} {[}\protect\hyperlink{ref-EcizztLg}{17}{]} is an open source toolbox for semantic segmentation, supporting unified benchmarking and state-of-the-art models ready to use out-of-box. It has become one of most widely used codebase for research in semantic segmentation (2.3K forks and 6.5K stars on GitHub as of September 29, 2023). This inspires us to develop \emph{MMV\_Im2Im} to facilitate research in image-to-image transformation with a special focus on biomedical applications.

First of all, different from general computer vision datasets, such as ImageNet {[}\protect\hyperlink{ref-lt4BNUoG}{18}{]}, where the images are usually small 2D RGB images (e.g., 3 x 256 x 256 pixels), biomedical applications usually involves large-scale high dimensional data (e.g., 500 images of 4 x 128 x 2048 x 2048 voxels). To deal with this issue, we employ the PersistentDataset in MONAI {[}\protect\hyperlink{ref-UU62HYC6}{19}{]} with partial loading and sampling support, as well as delayed image reading powered by aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{20}{]} as default (configurable if another dataloader is preferred). As a result, in our stress test, training a 3D nuclei instance segmentation model with more than 125,000 3D images can be conducted efficiently in a day, even with limited resources.

@@

Third, the toolbox pre-packs common functionalities specific to microscopy images. For example, we incorporate the special image normalization method introduced in {[}\protect\hyperlink{ref-Yq8wZ6hc}{5}{]}, where only the middle chunk along the Z dimension of 3D microscopy images will be used for calculating the mean and standard deviation of image intensity for standard normalization. Also, 3D light microscopy images are usually anisotropic, i.e., much lower resolution along Z than XY dimension. So, we adopt the anisotropic variation of UNet as proposed in {[}\protect\hyperlink{ref-jM3v1UjQ}{21}{]}.

Finally, to deploy the model in production, a model trained on small 3D patches sometimes needs to be applied not only on much larger images. Combining the efficient data handling of aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{20}{]} and the sliding window inference with gaussian weighted blending, the toolbox can yield efficient inference without visible stitching artifacts in production.

All in all, the \emph{MMV\_Im2Im} toolbox stands on the shoulders of many giants in the open-source software and ML engineering communities (pytorch-lightning, MONAI, aicsimageio, etc.) and is systematically designed for image-to-image transformation R\&D for biomedical applications. The source code of \emph{MMV\_Im2Im} is available at {[}\protect\hyperlink{ref-10dtMviwb}{1}{]}. This manuscript is generated with the open-source package Manubot {[}\protect\hyperlink{ref-YuJbg3zO}{22}{]}. The manuscript source code is available at {[}\protect\hyperlink{ref-B5mAMFLK}{23}{]}.

@@

The labelfree method refers to a DL method that can predict fluorescent images directly from transmitted light brightfield images {[}\protect\hyperlink{ref-Yq8wZ6hc}{5}{]}. Compared to brightfield images, fluorescent images can resolve subcellular structures in living cells at high resolution but with the cost of expensive and slow procedures and high phototoxicity. The labelfree method provides a new perspective in assay development to conduct integrated computational analysis of multiple organelles only with a single brightfield image acquisition. In our first demonstration, we applied \emph{MMV\_Im2Im} to build 2D/3D models that can predict fluorescent images of nuclear structures from brightfield images. For 3D models, we also compared (1) different image normalization methods, (2) different network backbones, and (3) different types of models.

It should be noted that while we recognize the importance of systematically evaluating the predictions, such an analysis falls outside the scope of this paper. We argue that an appropriate evaluation methodology should depend on specific downstream quantitative analysis goals (e.g., {[}\protect\hyperlink{ref-C2iqR6xE}{26}{]}). For example, if our aim is to quantify the size of nucleoli, we must compare the segmentation derived from real nucleoli signals to that of the predicted nucleoli segmentation, ensuring that measurements from both are consistent. Alternatively, if the goal is to localize the nucleoli roughly within the cell, Pearson correlation may be a more appropriate metric. In this work, we concentrate on visual inspection, using Pearson correlation and structural similarity as a rough quantitative reference. Our intent is to demonstrate the utility of our \emph{MMV\_Im2Im} package, and leave appropriate evaluations to users in their specific problems in real studies.

\emph{2D Labelfree:} We started with a simple problem using 2D images from the HeLa ``Kyoto'' cells dataset {[}\protect\hyperlink{ref-xv2VIyRP}{27}{]}. For all images, we took the brightfield channel and the mCherry-H2B channel out of the multi-channel timelapse movies. 2D images were acquired at 20x with 0.8 N.A. and then downscaled by 4 (pixel size: 0.299 nm x 0.299 nm). Example predictions can be found in Figure \ref{fig:labelfree_comparison}-A. We compared a basic UNet model {[}\protect\hyperlink{ref-TutLhFSz}{10}{]} and a 2D version of the fnet model in {[}\protect\hyperlink{ref-Yq8wZ6hc}{5}{]}. The fnet model achieved slightly more accurate predictions than the basic UNet, as seen in Figure \ref{fig:labelfree_comparison}-A.

\emph{3D Labelfree:} We tested with 3D images from the hiPSC single cell image dataset {[}\protect\hyperlink{ref-5sGcmDuy}{28}{]}. Specifically, we extracted the brightfield channel and the structure channel from the full field-of-view (FOV) multi-channel images, from the HIST1H2BJ, FBL, NPM1, LMNB1 cell lines, so as to predict from one brightfield image various nuclear structures, histones, nucleoli (dense fibrillar component via fibrillarin), nucleoli (granular component via nucleophosmin), and nuclear envelope, respectively. Images were acquired at 100x with 1.25 NA (voxel size: 0.108 micron x 0.108 micron x 0.29 micron).

@@

We conducted three groups of comparisons (see results in Figure \ref{fig:labelfree_comparison}-B). First, we compared three different image normalization methods for 3D images: percentile normalization, standard normalization, center normalization {[}\protect\hyperlink{ref-Yq8wZ6hc}{5}{]}. Percentile normalization refers to cutting the intensity out of the range of {[}0.5, 99.5{]} percentile of the image intensity and then rescale the values to the range of {[}-1, 1{]}, while the standard normalization is simply subtracting mean intensity and then divided by the standard deviation of all pixel intensities. Center normalization is similar to standard normalization, but the statistics are calculated only around the center along the Z-axis {[}\protect\hyperlink{ref-Yq8wZ6hc}{5}{]}. One could easily test different percentiles or rescaling to {[}0, 1{]} instead of {[}-1, 1{]}. Qualitatively, we found center normalization slightly more accurate and more robust than the other two (see first row of Figure \ref{fig:labelfree_all_structures}-B).

Second, we compared different network backbone architectures, including the original fnet model {[}\protect\hyperlink{ref-Yq8wZ6hc}{5}{]}, an enhanced UNet {[}\protect\hyperlink{ref-M7480NLD}{29}{]}, the attention UNet {[}\protect\hyperlink{ref-OCow1hly}{30}{]}, two transformer-based models, SwinUNETR {[}\protect\hyperlink{ref-ZWL3IrVc}{31}{]} and UNETR{[}\protect\hyperlink{ref-XCKUntOB}{32}{]} (all with center normalization). Inspecting the predictions on a holdout validation set suggested that fnet achieved the best performance.

@@

Segmentation is a common image processing task, and can be considered as a special type of image-to-image transformation, where the generated images are segmentation masks. DL-based methods have achieved huge success in semantic segmentation in biomedical images. In this example, we demonstrated \emph{MMV\_Im2Im} on a pathology application to segment glands from hematoxylin and eosin (H\&E) stained tissue images from the 2015 Gland Segmentation challenge {[}\protect\hyperlink{ref-45Sirz1X}{33},\protect\hyperlink{ref-XAffSYIR}{34}{]}. Stain normalization is an important pre-processing step in order to develop models robust to stain variation and tissue variations. \emph{MMV\_Im2Im} included a classic stain normalization method {[}\protect\hyperlink{ref-tQhnZyjK}{35}{]} as a pre-processing step. The effect of stain normalization can be observed in Figure \ref{fig:2d_gland}-A and B. We trained a simple attention UNet model {[}\protect\hyperlink{ref-OCow1hly}{30}{]}. Evaluated on the two different holdout test sets, the model achieved F1-score, 0.904 ± 0.060 and 0.861 ± 0.117 on test set A and test set B, respectively. The performance was competitive compared to the methods reported in the challenge report {[}\protect\hyperlink{ref-XAffSYIR}{34}{]}, especially with much more consistent performance across the two different test sets. Example results can be found in Figure \ref{fig:2d_gland}-C.

@@

Instance segmentation is a type of segmentation problem that goes beyond semantic segmentation. The goal is to differentiate not only between different types of objects, but also different instances of the same type of objects. Currently, the \emph{MMV\_Im2Im} package supports EmbedSeg-type models. The major benefit of EmbedSeg-type models is their agnosticism to the morphology and dimensionality of the object instances, compared to other models such as StarDist {[}\protect\hyperlink{ref-tIIG2f8K}{36},\protect\hyperlink{ref-14h90Vfg0}{37}{]}, SplineDist {[}\protect\hyperlink{ref-17Yrl6WGQ}{38}{]} and Cellpose {[}\protect\hyperlink{ref-TugPkOLy}{39}{]}. For example, different from the others, EmbedSeg-type models are even able to generate instance segmentation where each instance contains multiple connected components. Additional frameworks such as Omnipose {[}\protect\hyperlink{ref-lXzmjM5n}{40}{]} will be supported in future versions. Another mainstream category of instance segmentation methods are detection-based models, such as Mask-RCNN {[}\protect\hyperlink{ref-xi8wnibR}{41}{]}. However, these models are better suited to the detection framework rather than image-to-image transformation (see Discussion section for details).

The \emph{EmbedSeg}-type models were re-implemented according to the original paper {[}\protect\hyperlink{ref-K2ugNcVa}{11},\protect\hyperlink{ref-QmYuUQ5K}{12}{]} following the generic boilerplate in \emph{MMV\_Im2Im}, with significant improvement. First of all, following the modular design of \emph{MMV\_Im2Im}, it is flexible to use different neural network models as the backbone. For 3D anisotropic microscopy images, the original backbone ERFNet {[}\protect\hyperlink{ref-XAkgs3Nh}{42}{]} doesn't take the anisotropic dimensions into account and therefore may not perform well or even be applicable. In this scenario, it is straightforward to employ another anisotropic neural network bone, such as the anisotropic U-Net in {[}\protect\hyperlink{ref-jM3v1UjQ}{21}{]} or the anisotropic version of Dynamic U-Net in MONAI. Second, we significantly improve training strategy. The original version requires pre-cropping patches centered around each instance and pre-calculated the center images and class images. This may generate a massive amount of additional data on the disk. More importantly, such pre-cropping makes data augmentation nearly impossible, except the simple ones like flipping (otherwise, the pre-calculated centers might be wrong), and also greatly undersamples around negative cases (e.g., background). For example, we have observed that for an EmbedSeg model training only with patches centered around instances, the model may suffer from degraded performance during inference when there are a large amount of background areas without any instances. Again, following the modular design of \emph{MMV\_Im2Im}, it is now possible to do on-the-fly data augmentation and patch sampling, even weighted patch sampling. Third, our improved \emph{EmbedSeg}-type models can accept an exclusion mask so that certain parts of the images can be ignored during training. This is especially useful for partially annotated ground truth. For large images, it could be extremely time-consuming to require every single instance to be annotated. The exclusion masks can address this bottleneck. Another extension compared to the original implementation was that the \emph{MMV\_Im2Im} package made sliding windowing inference straightforward, and therefore permitted easy handling of images of any size during inference.

In this work, we tested on both 2D and 3D instance segmentation problems. Going from 2D to 3D is not a simple generalization from 2D models by switching 2D operations with 3D operations, but with many practical challenges. Large GPU footprint is one of the biggest issues, which makes many training strategies common in 2D not feasible in 3D, e.g.~limited mini-batch size. \emph{MMV\_Im2Im} is able to take advantage of state-of-the-art ML engineering methods to efficiently handle 3D problems. For example, by using effective half-precision training, one can greatly reduce GPU memory workload for each sample and therefore increase the batch size or the patch size. When multiple GPUs are available, it is also possible to easily take advantage of the additional resources to scale up the training to multiple GPU cards, even multiple GPU nodes. As a demonstration, we applied \emph{EmbedSeg}-like models to a 2D problem of segmenting \emph{C. elegans} from widefield images {[}\protect\hyperlink{ref-138foKNOh}{43}{]}, as well as a 3D problem of nuclear segmentation from fluorescent and brightfield images from the hiPSC single cell image dataset {[}\protect\hyperlink{ref-5sGcmDuy}{28}{]}.
